Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-22 paper/paper.tex
intro: gale-et-al review, some motivation for cplx-vd

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embeded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and numerical computational
compelexity restrictions.

These applications establish a need for sparsification and model compression techniques.
However audio or radio frequiency signal analysis is inherently complex-valued, yet
$\cplx$-valued models and neural networks and, in particular, comprssion methods tailored
to $\cplx$-networks remain a niche field of research. Although computational efficiency
reduction methods such as quantization and integer-based arithmetic appear to be directly
transferrable to $\cplx$-field, \citep{citation_needed}, the most promising methods,
\citep{gale_state_2019} such as weight pruning \citep{zuo_compression_2019}, bayesian 
dropout \citep{kingma_variational_2015}, and \citep{louizos_learning_2017} require
additional considerations.

% on sparsification
Solving the storage and compute constraints encourages development of model compression
and sparsification methods, that focus on favourable tradeoff between performance and size.
%
Most popular techniques include
model quantization from IEEE754,
and dropout. The key assumption is that the more explicitly zero parameters there is,
the less computations is needed.

@@

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, thereby breaking up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propse
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entroy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possiblity of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learnig capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}

% segue into bayesian inference
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs accoring to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian inference technique. Bayesian Inference as a principled framework of reasoning
about uncertainty, updates prior beliefs in accordance with the likelihood of empirical
observations into posterior belief. Under suitable modelling and prior assumptions, Bayesian
methods can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

>>>>

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and numerical computational
complexity restrictions. Therefore compression methods for achieving high model sparsity
and numerical efficiency without losing much in performance are especially relevant.

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research. Computational efficiency
reduction methods such as quantization, integer-based arithmetic \citep{citation_needed},
and distillation \citep{distillation} appear to be immediately applicable to $\cplx$-field.
%
Some model sparsification methods, e.g. magnitude pruning \citep{zuo_compression_2019}, which
reactivates zeroed parameters based on gradients, and sparsity inducing regularizers, can be
adapted directly to $\cplx$-valued models. Other methods -- require additional considerations.

\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015}, $\ell_0$
regularization \citep{louizos_learning_2017} and magnitude pruning on large-scale models.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning on the studied architectures, overall pruning is preferable due to
simplicity and speed. They also conclude that, although dropout and $\ell_0$ yield less
consistent performance and are more complex, they induce non-uniform sparsity throughout
the model, which is essential for better compression \verify{(He et al., 2018)}.
%
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%
Therefore it appears that Bayesian and stochastic sparsification techniques can be useful
for models in $\cplx$ domain. \verify{$\ell_0$ -- stoch?}

% on sparsification
Solving the storage and compute constraints encourages development of model compression
and sparsification methods, that focus on favourable tradeoff between performance and size.
%
Most popular techniques include
model quantization from IEEE754,
and dropout.

The key assumption is that the more explicitly zero parameters there is,
the less computations is needed.

@@

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, thereby breaking up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}

% segue into bayesian inference
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian inference technique. Bayesian Inference as a principled framework of reasoning
about uncertainty, updates prior beliefs in accordance with the likelihood of empirical
observations into posterior belief. Under suitable modelling and prior assumptions, Bayesian
methods can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.
